---
title: "Using Machine Learning to Quantify Performance in Exercise"
author: "Andrew Hwang"
output: html_document
---

## Introduction

In this project, we will look at quantifying performance in exercise. We will use data collected in a [study](http://groupware.les.inf.puc-rio.br/har) where test subjects wore accelerometers and performed various exercises bother properly and improperly. We will use random forests, a machine learning algorithm, to predict whether or not a person is performing an exercise correctly.

## Obtaining the Data

The training data for this project is available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv).

```{r}
if (!file.exists("./pml-training.csv")) {
    download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "./pml-training.csv", method = "curl")
}
```

## Data Processing

We will first import the data and get some information about the dataset.

```{r}
data = read.csv("./pml-training.csv", na.strings = c("NA",""))
str(data)
summary(data$classe)
```

From this, we can see that there are five possible classes for the variable we are predicting on, which is *classe*. Next, we will split the data into a training and testing set.

```{r}
set.seed(100)
library(caret)
inTrain = createDataPartition(y=data$classe, p=0.7, list=FALSE)
training = data[inTrain,]
testing = data[-inTrain,]
```

Next, we will clean the dataset by removing variables that have many missing values.

```{r}
na_test = sapply(training, function (x) {sum(is.na(x))})
table(na_test)
```

From this, we see that 100 columns have missing values. We will remove these columns and use the remaining 60.

```{r}
empty_cols = names(na_test[na_test == 13443])
training = training[, !names(training) %in% empty_cols]
str(training)
```

Since we are only interested in variables that have to do with movement, we will remove the first seven variables that are user info and timestamps.

```{r}
training = training[,-c(1:7)]
```

## Creating the Model

We will be using random forests to build the model, as it is among the most accurate machine learning algorithms.

```{r, cache=TRUE}
library(caret)
library(randomForest)
if (!file.exists("./rfmodel.RDS")) {
   model = train(classe~., method="rf", data=training)
   saveRDS(model, "rfmodel.RDS") 
}
model = readRDS("./rfmodel.RDS")
```

## Evaluating the Model

We will now use the model built from the training dataset and apply it to the testing dataset to see how accurate it is. The accuracy is defined as the percentage of correct predictions.

```{r}
mean(predict(model, testing) == testing$classe)*100
```

## Results

The model was 99.32% accurate, so the error is .68%. This means the model provides accurate predictions, and can be used to determine which of the five movements the person is performing.